---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# cranlift: Serve CRAN-like Repositories as RESTful APIs

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/cranlift)](https://cran.r-project.org/package=cranlift)
<!-- badges: end -->

Like the `miniCRAN` and `drat` packages, `cranlift` allows users to manage their
own CRAN-like R package archives. However, `cranlift` does so by running a
fully-featured API server that exposes packages as resources that can be added,
deleted, and modified using HTTP requests, in line with RESTful design
principles.

`cranlift` supports both source and binary packages, as well as CRAN's Archive
layout for storing old versions of packages.

## Installation

`cranlift` is not yet available on [CRAN](https://CRAN.R-project.org). For now,
you can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("atheriel/cranlift")
```

## Usage

To serve an (empty) CRAN-like repository, call `cranlift::serve()`:

```{r example-serve, eval = FALSE}
dir.create(repo <- tempfile("repo"))
cranlift::serve(repo, port = 8000)
```
```{r bg-serve, echo = FALSE}
p <- callr::r_bg(function() {
  dir.create(repo <- tempfile("repo"))
  cranlift::serve(repo, port = 8000)
}, supervise = TRUE)
stopifnot(p$is_alive())
```

You can then verify that R sees it as a package repository (with zero packages
at the moment):

```{r example-available}
nrow(available.packages(repos = "http://127.0.0.1:8000"))
```

Now let's add an example package to the repository, say `miniCRAN`. Because
we're interacting with a server, we do this with an HTTP request:

```{r example-add}
path <- download.packages("miniCRAN", tempdir(), type = "source")[1,2]
httr::PUT(
  sprintf("127.0.0.1:8000/src/contrib/%s", basename(path)),
  body = list(file = httr::upload_file(path))
)

nrow(available.packages(
  repos = "http://127.0.0.1:8000", ignore_repo_cache = TRUE
))
```

Deleting a package (say, because a new version is available) works similarly:

```{r example-delete}
httr::DELETE(
  sprintf("127.0.0.1:8000/src/contrib/%s", basename(path))
)

nrow(available.packages(
  repos = "http://127.0.0.1:8000", ignore_repo_cache = TRUE
))
```

By default, deleted packages will be moved to the archive:

```{r example-archive}
httr::HEAD(sprintf(
  "127.0.0.1:8000/src/contrib/Archive/miniCRAN/%s", basename(path)
))
```

## Usage with `drat` and `miniCRAN`

`cranlift` can serve repositories created with `drat` and `miniCRAN` (and has a
strong emphasis on compatibility with the common repository layout they use),
but be warned that it may modify the `PACKAGES` files these packages create.

## License

`cranlift` is made available under the terms of the MIT license. See the
`LICENSE.md` file for details.
